---
output: github_document
---

# usdarnass

```{r setup, include = FALSE}
library("knitr")
library("usdarnass")
opts_chunk$set(collapse = TRUE,
               comment = "#>",
               fig.path = "man/figures/README-",
               out.width = "100%")
```

[![Travis-CI Build Status](https://travis-ci.org/rdinter/usdarnass.svg?branch=master)](https://travis-ci.org/rdinter/usdarnass)

An alternative for downloading various USDA data from [https://quickstats.nass.usda.gov/](https://quickstats.nass.usda.gov/) through R. You must sign up for an [API key](https://quickstats.nass.usda.gov/api) from the mentioned website in order for this package to work. Disclaimer:

> This product uses the NASS API but is not endorsed or certified by NASS.


# Install

Eventually, this package will be uploaded to [CRAN](https://cran.r-project.org/). At which point, you should be able to install the package through the typical method:

```{r installation_cran, eval = FALSE}
install.packages("usdarnass")
```

However, until the package has been submitted to CRAN the only way to use the package is through the development version (needs devtools installed):

```{r installation_github, eval = FALSE}
# install.packages("devtools")
devtools::install_github("rdinter/usdarnass")
```

Once installed through either method, you can load package the conventional way:

```{r load}
library("usdarnass")
```


# Usage

See the [Get started](articles/usdarnass.html) article for a more in depth discussion of the package features.


# Documentation

See the [documentation site](http://rdinter.github.io/usdarnass/).


# See Also

There are other `R` packages that work with USDA Quick Stats that may be more useful to others:

* [emraher/rnass](https://github.com/emraher/rnass)
* [potterzot/rnassqs](https://github.com/potterzot/rnassqs)